---
title: "Computational Modeling - Assignment 1"
author: "Riccardo Fusaroli"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. this markdown has 2 parts as it spans 2 weeks of teaching

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't, peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
# we define a density of 20:
dens = 20

# We create a grid of possible p's from 0 to 1:
p_grid <- seq(from = 0 , to = 1 , length.out = dens)


# We define a prior that is centered around chance:
prior <- dnorm(p_grid, 0.5, 0.1)

# We compute the likelihood for all p's:
likelihood <- dbinom( 3 , size = 6 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)


# We make a plot
# (First we take all our lists and put them in a data frame)

d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

library(ggplot2)

ggplot(d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Riccardo's Knowledge of CogSci")+ ylab("posterior probability")


# For the quadratic approximation we need
library(rethinking)

# We create a quap with C (correct) and I (incorrect) following a binomial likelihood distribution and has a uniform prior
riccardo.qa <- quap(
  alist(
    C ~ dbinom( C + I ,p) , # binomial likelihood
    p ~ dunif(0,1) # uniform prior
  ) ,
  data=list(C=3,I=3) )

# We get a summay'ry of the quap
precis(riccardo.qa)


# analytical calculation
C <- 3
I <- 3
curve( dbeta( x , C+1 , I+1 ) , from=0 , to=1 )

# quadratic approximation
curve( dnorm( x , 0.5 , 0.2 ) , lty=2 , add=TRUE )



# What is the probability that Riccardo will get more than 0.5 of the question correct?
# We need to sum all the plausabilities for the different p's that are above 0.5
sum(posterior[p_grid > 0.5] )

# The probability of getting

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.


```{r}
dens <- 20
p_grid <- seq(from = 0 , to = 1 , length.out = dens)

#Define the prior
prior <- dnorm(p_grid, 0.5, 0.1) 

#Compute the likelihood at each value in grid
likelihood <- dbinom( 2 , size = 2 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

#Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

#Draw the plot

Kristian_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

Kristian <- ggplot(Kristian_d, aes(grid, posterior)) +  geom_point() +geom_line()+theme_classic()+geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's Knowledge of CogSci")+ ylab("posterior probability")

Kristian


####################################################################################

likelihood <- dbinom(160 , size = 198 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

#Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

Daina_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

Daina <- ggplot(Daina_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Dains's Knowledge of CogSci")+ ylab("posterior probability")

Daina

####################################################################################

likelihood <- dbinom( 66 , size = 132 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior <- likelihood * prior

#Standardize the posterior (so it sums to 1)
posterior <- unstd.posterior / sum(unstd.posterior)

Mikkel_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

Mikkel <- ggplot(Mikkel_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Mikkel's Knowledge of CogSci")+ ylab("posterior probability")

Mikkel



####################################################################################

# 2.a

# We have already plottet posteriors for them all, so here we just need the prior plots:

# We will use the "sensibly centered at chance" prior for each teacher

# Riccardo:

dens(rbinom(1e4, 6, rnorm(1e4, 0.5, 0.1)))  # A sample from a binomial distribution, where 10,000 values are randomly sampled from a normal distribution with a mean of 0.5 and a sd of 0.1

# Kristian:

dens(rbinom(1e4, 2, rnorm(1e4, 0.5, 0.1))) 

# Daina:

dens(rbinom(1e4, 198, rnorm(1e4, 0.5, 0.1)))

# Mikkel:

dens(rbinom(1e4, 132, rnorm(1e4, 0.5, 0.1)))


```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
# We change the prior
prior <- dnorm(p_grid, 0.8, 0.2)

###############################################################

# Riccardo

# Prior plot
dens(rbinom(1e4, 6, rnorm(1e4, 0.8, 0.2)))


# We compute the likelihood for all p's:
likelihood <- dbinom( 3 , size = 6 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# saving as df
Riccardo_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior probability
ggplot(Riccardo_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Riccardo's Knowledge of CogSci")+ ylab("posterior probability")


###############################################################

# Kristian

# Prior plot
dens(rbinom(1e4, 2, rnorm(1e4, 0.8, 0.2)))


# We compute the likelihood for all p's:
likelihood <- dbinom( 2 , size = 2 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Kristian_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Kristian_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's Knowledge of CogSci")+ ylab("posterior probability")



##############################################################

# Daina

# Prior plot
dens(rbinom(1e4, 198, rnorm(1e4, 0.8, 0.2)))


# We compute the likelihood for all p's:
likelihood <- dbinom( 160 , size = 198 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Daina_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Daina_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Daina's's Knowledge of CogSci")+ ylab("posterior probability")




##############################################################

# Mikkel

# Prior plot
dens(rbinom(1e4, 132, rnorm(1e4, 0.8, 0.2)))


# We compute the likelihood for all p's:
likelihood <- dbinom( 66 , size = 132 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Mikkel_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Mikkel_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Mikkels's Knowledge of CogSci")+ ylab("posterior probability")









```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?


First with a flat (uniform) prior:
```{r}
# We change the prior
prior <- rep(1, dens) # FLAT

###############################################################

# Riccardo


# We compute the likelihood for all p's:
likelihood <- dbinom( 300 , size = 600 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# saving as df
Riccardo_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior probability
ggplot(Riccardo_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Riccardo's Knowledge of CogSci")+ ylab("posterior probability")


###############################################################

# Kristian



# We compute the likelihood for all p's:
likelihood <- dbinom( 200 , size = 200 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Kristian_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Kristian_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristians's Knowledge of CogSci")+ ylab("posterior probability")



##############################################################

# Daina



# We compute the likelihood for all p's:
likelihood <- dbinom( 16000 , size = 19800 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Daina_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Daina_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Daina's Knowledge of CogSci")+ ylab("posterior probability")




##############################################################

# Mikkel



# We compute the likelihood for all p's:
likelihood <- dbinom( 6600 , size = 13200 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Mikkel_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Mikkel_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Mikkels's Knowledge of CogSci")+ ylab("posterior probability")



```


And now with the normally distributed prior around 0.8 with a sd of 0.2:

```{r}
# We change the prior
prior <- dnorm(p_grid, 0.8, 0.2) 

###############################################################

# Riccardo


# We compute the likelihood for all p's:
likelihood <- dbinom( 300 , size = 600 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# saving as df
Riccardo_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior probability
ggplot(Riccardo_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Riccardo's Knowledge of CogSci")+ ylab("posterior probability")


###############################################################

# Kristian



# We compute the likelihood for all p's:
likelihood <- dbinom( 200 , size = 200 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Kristians data as a df
Kristian_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Kristian_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristians's Knowledge of CogSci")+ ylab("posterior probability")



##############################################################

# Daina



# We compute the likelihood for all p's:
likelihood <- dbinom( 16000 , size = 19800 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Daina's data as a df
Daina_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Daina_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Daina's Knowledge of CogSci")+ ylab("posterior probability")




##############################################################

# Mikkel



# We compute the likelihood for all p's:
likelihood <- dbinom( 6600 , size = 13200 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
posterior <- unstd.posterior / sum(unstd.posterior)

# Saving Mikkels data as a df
Mikkel_d <- data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# plotting posterior prob
ggplot(Mikkel_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Mikkels's Knowledge of CogSci")+ ylab("posterior probability")

```





5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?


Answered in the report.

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)? 

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Daina: 148 correct answers out of 172 questions (again, Daina never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)



```{r}
# We start getting the original posteriors for each teacher (using the data that is not scaled up 100 times)

# We set the prior
prior <- dnorm(p_grid, 0.8, 0.2)

###############################################################

# Riccardo


# We compute the likelihood for all p's:
likelihood <- dbinom( 3 , size = 6 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
R_posterior <- unstd.posterior / sum(unstd.posterior)


###############################################################

# Kristian


# We compute the likelihood for all p's:
likelihood <- dbinom( 2 , size = 2 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
K_posterior <- unstd.posterior / sum(unstd.posterior)


##############################################################

# Daina


# We compute the likelihood for all p's:
likelihood <- dbinom( 160 , size = 198 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
D_posterior <- unstd.posterior / sum(unstd.posterior)



##############################################################

# Mikkel


# We compute the likelihood for all p's:
likelihood <- dbinom( 66 , size = 132 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * prior


# We standardize the posterior:
M_posterior <- unstd.posterior / sum(unstd.posterior)



#################################################################


# Now we are ready to make the new models for each teacher using the old posteriors as priors:


#################################################################


# Riccardo

# We set the new prior to be the old posterior
R_prior <- R_posterior

# We compute the likelihood for all p's:
likelihood <- dbinom( 9 , size = 10 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * R_prior


# We standardize the posterior:
R_posterior <- unstd.posterior / sum(unstd.posterior)


# saving as df
Riccardo_d <- data.frame(grid = p_grid, posterior = R_posterior, prior = R_prior, likelihood = likelihood)


# plotting posterior prob
ggplot(Riccardo_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'red')+  xlab("Riccordo's Knowledge of CogSci")+ ylab("posterior probability")



#################################################################


# Kristian


# We set the new prior to be the old posterior
K_prior <- K_posterior


# We compute the likelihood for all p's:
likelihood <- dbinom( 8 , size = 12 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * K_prior


# We standardize the posterior:
K_posterior <- unstd.posterior / sum(unstd.posterior)



# saving as df
Kristian_d <- data.frame(grid = p_grid, posterior = K_posterior, prior = K_prior, likelihood = likelihood)


# plotting posterior prob
ggplot(Kristian_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'red')+  xlab("Kristian's Knowledge of CogSci")+ ylab("posterior probability")



##################################################################

# Daina


# We set the new prior to be the old posterior
D_prior <- D_posterior



# We compute the likelihood for all p's:
likelihood <- dbinom( 148 , size = 172 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * D_prior


# We standardize the posterior:
D_posterior <- unstd.posterior / sum(unstd.posterior)




# saving as df
Daina_d <- data.frame(grid = p_grid, posterior = D_posterior, prior = D_prior, likelihood = likelihood)


# plotting posterior prob
ggplot(Daina_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'red')+  xlab("Daina's Knowledge of CogSci")+ ylab("posterior probability")


###################################################################

# Mikkel


# We set the new prior to be the old posterior
M_prior <- M_posterior


# We compute the likelihood for all p's:
likelihood <- dbinom( 34 , size = 65 , prob = p_grid )


# We compute an unstandarized posterior:
unstd.posterior <- likelihood * M_prior


# We standardize the posterior:
M_posterior <- unstd.posterior / sum(unstd.posterior)


# saving as df
Mikkel_d <- data.frame(grid = p_grid, posterior = M_posterior, prior = M_prior, likelihood = likelihood)


# plotting posterior prob
ggplot(Mikkel_d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'red')+  xlab("Mikkels's Knowledge of CogSci")+ ylab("posterior probability")


```


This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Daina: 148 correct answers out of 172 questions (again, Daina never gets bored)
- Mikkel: 34 correct answers out of 65 questions

